One-stage point transect distance sampling using iterated integrated nested Laplace approximations

### 2024-10-30 paper.tex
update pairwise distances

To evaluate the ability of the SPDE effect to capture the spatial heterogeneity in the observed data we generated 1,000 point patterns using the posterior intensity field and thinned them with the posterior detection function to create 1,000 simulated datasets to compare with the observed data.  We calculated the pairwise distances between all observations for each of these datasets and compared the frequency of pairwise distances with those in the observed data.
\begin{figure}[!htb]
	\begin{center}
		\includegraphics[scale=0.6]{figures/post_pp_distances.png}
		\caption{Boxplots of the frequency of pairwise distances (tick mark values for end of each bin) based on 1,000 simulated datasets.  Red circles indicate the frequency of pairwise distances in the observed data.}
		\label{fig:post-pp-distances}
	\end{center}
\end{figure}
Figure \ref{fig:post-pp-distances} shows that the SPDE effect broadly captures the spatial clustering in the observed data.  There is some evidence of under-predicting the strength of the clustering at shorter distances between 1,000 and 2,500 m, with the simulated data showing some evidence of negative bias.  This could be due to over-smoothing in some areas which can happen if using a stationary model.  Large areas of the study region having low density, which has a larger correlation range than higher density areas, and so the effect will average between these.

>>>>

To evaluate the ability of the SPDE effect to capture the spatial heterogeneity in the observed data we generated 1,000 point patterns using the posterior intensity field and thinned them with the posterior detection function to create 1,000 simulated datasets to compare with the observed data.  We calculated the pairwise distances between all observations for each of these datasets and compared the frequency of pairwise distances with those in the observed data.
\begin{figure}[!htb]
	\begin{center}
		\includegraphics[scale=0.6]{figures/post_pp_distances.png}
		\caption{Boxplots of the frequency of pairwise distances (tick mark values for end of each bin) based on 500 simulated datasets from the posterior predictive distribution.  Red circles indicate the frequency of pairwise distances in the observed data.}
		\label{fig:post-pp-distances}
	\end{center}
\end{figure}
Figure \ref{fig:post-pp-distances} shows that the SPDE effect broadly captures the spatial clustering in the observed data.  There is some evidence of under-predicting the strength of the clustering at shorter distances between 1,000 and 2,500 m, with the simulated data showing some evidence of negative bias.  This could be due to over-smoothing in some areas which can happen if using a stationary model.  Large areas of the study region having low density, which has a larger correlation range than higher density areas, and so the effect will average between these.
